---
title: "DC_Mar26 Instructor Workshop"
author: "Sabah Ul-Hasan"
date: "3/26/2018"
output: html_document
---
Interested in becoming a Software or Data Carpentry Instructor, but unsure if you're qualified? Here's a breakdown of the range of experiences from participants Instructor Workshop Mar 26-27, 2018 at the UC Davis DIB (http://dsi.ucdavis.edu/) demonstrating people of all levels and backgrounds can all start as excellent instructors! 

Step 1: Creating a data frame in R
```{r}
# Create lists for each question and associated answer
Q1_WorkshopParticipation = c( "Taken a Workshop", "Workshop Helper", "Organized a Workshop", "No BUT Familiar", "No NOT Familiar" )
A1_WorkshopParticipation = c(7,9,0,11,2)

Q2_TeachingLevel = c("UG or Grad Assistant Experience", "No Experience", "Taught Seminar or Workshop", "Instructor of Record", "Primary Education Experience", "Outreach or Demo Experience")
A2_TeachingLevel = c(15,0,14,11,3,7)

# Join the lists into a vector of equal lengths
Q1_df = data.frame(Q1_WorkshopParticipation, A1_WorkshopParticipation)
Q2_df = data.frame(Q2_TeachingLevel, A2_TeachingLevel)
```

Step 2: Visualizing the results of each dataset
```{r}
# Install appropriate packages
install.packages(ggplot2)
library(ggplot2)
packageDescription("ggplot2") # v2.2.1
 
# Simple barplot of Question 1, "Have you ever participated in a Software Carpentry or Data Carpentry Workshop?"
Q1 <-
  ggplot(data = Q1_df, aes(x=Q1_WorkshopParticipation, y=A1_WorkshopParticipation)) + # plot
  xlab("Level of Previous Participation") + ylab("Number of Participants") + # labeling axes
  geom_bar(stat="identity", color="steel blue", fill="gold") + # aesthetic
  ggtitle("Question 1: Participation in a SW or Data Carpentry Workshop") + # maine title
  geom_text(aes(label=A1_WorkshopParticipation), vjust=2, color="steel blue", size=5) # see number

Q1 # view plot

# Simple barplot of Question 2, "Which of these most accurately describes your teaching experience?"
Q2 <-
  ggplot(data = Q2_df, aes(x=Q2_TeachingLevel, y=A2_TeachingLevel)) + # plot
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) + # rotate x axis labels
  xlab("Level of Previous Teaching Experience") + ylab("Number of Participants") + # labeling axes
  geom_bar(stat="identity", color="gold", fill="steel blue") + # aesthetic
  ggtitle("Question 2: Description of Teaching Experience") + # maine title
  geom_text(aes(label=A2_TeachingLevel), vjust=1.5, color="gold", size=5) # see number

Q2 # view plot
```

Step 3: Creating and saving a figure from the results
```{r}
# Install appropriate packages
install.packages(gridExtra)
library(gridExtra)
packageDescription("gridExtra") # v2.3

png("Figure1.png") # save as a png
grid.arrange(Q1, Q2) # view plots on same page
dev.off() # save to your working directory
```

Woohoo, you did it!