---
title: "DataCarpentryWorkshop_File1"
author: "Sabah Ul-Hasan"
date: "11/1/2018"
output:
  html_document: default
---

Nov 1-2, 2018 UC Merced Workshop Schedule (by Nora C Mitchell, PhD)
https://noracmitchell.github.io/2018-11-01-ucmerced/

Etherpad
https://pad.carpentries.org/_2018-11-01-ucmerced

Unix  http://man7.org/linux/man-pages/man1/ls.1.html 
      https://www.shell-tools.net/index.php?op=regexp
      https://regex101.com/r/yE3tO8/1
  Best practices: avoid spaces and underscores in file folder names
  Text editors handle text. nano is a text editor. 
    In nano, the ^ symbol means to press Control key.
  Microsoft Word is not a text editor, it's a word processor.
  Reuse a previously entered command by pressing the "up arrow"
  Pressing tab helps fill in the rest of your file name in a command
  Once space is the same as multiple spaces
  No space vs one space results in different outcomes 
  pwd   = present working directory
  cd    = change directory
  cd .. = moves you up in directory
  cd -  = takes you back to your previous working directory
  ls    = shows what's in your directory
  mkdir = makes a new directory
  rm    = removes a file in your directory (cannot recover)
  rm -r = remove folder
  mv    = moves a file, changes its name (overwrites file)
  cp    = copies a file
  cp -r = copies a directory
  wc    = gives you the line, word, and character count of a file
  wc -c = character count
  wc -l = line count
  wc -w = word count
  To make a new file, type:           > myfile.txt   
  To open that file with nano, type:  nano myfile.txt 
  Close the file by pressing:         CTRL + X
  cat draft.txt will show you the text in draft.txt
  nano /draft.txt will open a new draft in nano named draft.txt 
  echo hello > testfile01.txt will write hello into testfile01.txt 
  echo hello >> testfile02.txt will append hello into testfile02.txt
  head -3 animals.txt will return the top three lines of animals.txt
  tail -2 animals.txt will return the bottom two lines of animals.txt


R     https://tinyurl.com/ycvyop3r 
      http://swcarpentry.github.io/r-novice-gapminder/
Data management       http://openrefine.org/
  project management  https://tinyurl.com/y77tbloo 
  jupyter             https://tinyurl.com/ybx6mbjm
  tidyverse           https://www.tidyverse.org/
                      http://hackmd.io/ 
  r-bloggers          https://tinyurl.com/ybfb9rqo 
Data vizualization    https://www.r-graph-gallery.com/
  ggplot2             https://tinyurl.com/ya4lxpfg 
  colorbrewer         https://tinyurl.com/ycklsjkf 
Code troubleshotting  https://stackoverflow.com/ 
                      https://www.r-bloggers.com/
                      
  Reading or writing format .csv, .txt
  Seeing code for function  print(function_name)
                            print(summary)
  Packages    Type function_name without brackets for function's code
              http://masterr.org/r/how-to-view-source-code-in-r/
    versions  https://tinyurl.com/ya25e85b       
  methods() View different methods
  Using R for large datasets and cluster vs RStudio for quick viz
  Lists   One-dimensional arrays of data containing multiple data types
  Vectors One-dimensional arrays of data with single data type
  Dataframe usually has rows that are lists and columns are vectors
    Rows are usually samples of a sort
    Columns are usually the observations


Git   https://noracmitchell.github.io/git-novice.html 
git init                  initiates a new repository
git add <filename>        stages that file / those files
git commit -m "message"   commits the staged files to the repository
git diff                  looks at diffs versions 
git log                   shows the commits
git log --oneline         shows commits using one line
git reset HEAD <filename> unstages the file
git status                status of the repository
git checkout              restore past state


Programming with R Lesson
0. Setup: Complete

1. Analyzing Patient Data
```{r setup, include=FALSE}
# Set working directory
setwd("~/Desktop/r-novice-inflammation/")
# Load data
read.csv(file = "data/inflammation-01.csv", header = FALSE)


#####  Variables  #####
# Assiging a value to a string
# Defining a variable: stores a value
weight_kg <- 55
weight_kg # Assigned 55 kg

# Convert kg to pounds (1 kg = 2.2 lbs)
2.2 * weight_kg
2.2 * 55 # Confirms that we correctly assigned our string to 55 kg
weight_lb <- 2.2 * weight_kg

rm(weight_lb) # This does the same function as manually removing with the broom in environment

weight_lb <- 2.2 * weight_kg

weight_kg <- 100.0
# weight in kg now...
weight_kg

weight_lb # these lbs are from an earlier, different value (55 kg)
#####  Variables  #####


# Printing with Parentheses
dat <- read.csv(file = "data/inflammation-01.csv", header = FALSE)
head(dat) # function head() lets us see first 6 lines of data frame

# Assigning Values to Variables
mass <- 47.5
age <- 122
mass <- mass * 2.0 # converts original mass to 95
age <- age - 20 # converts original age to 102

# Manipulating Data
class(dat) # what's the classifcation?
class(x)
apples <- "oranges"
class(apples)

dim(dat) # whats the dimension
dat[1 , 1] # tells us what's in the first row of the first column [ row , column]
dat[30, 20]

subdata <- dat[c(1, 3, 5), c(10, 20)] # subsets data

1:5 # list of 1 through 5
3:12 # list of numbers 3 to 12
dat[1:4, 1:10] # gives output of rows 1-4 x columns 1-10
dat[5:10, 1:10] # gives output for rows 5-10 x col 1-10
dat[5, ] # every value for all columns with row 5
dat[, 16:18] # all values for columns 16-18

# Addressing Columns by Name
patient_1 <- dat[1, ] # subset to look at patient 1
max(patient_1) # max() finds maximum value [inflammation]

max(dat[2, ]) # subset for patient 2 and find max

min(dat[, 7]) # column 7 what was smallest inflammation
# v values are corresponding to days (40 days)
mean(dat[, 7]) # on day 7 the mean was 3.8 inflammation
median(dat[, 7]) # on day 7 the median was 4
sd(dat[, 7]) # on day 7 standard dev was 1.73

# Forcing Conversion
summary(dat[, 1:4]) # summary of data for days 1-4
?summary() # question mark in front tells us what it is

?apply() # what does apply do?
avg_patient_inflammation <- apply(dat, 1, mean)
# identifies the mean in each patient and creates list
class(avg_patient_inflammation)
avg_day_inflammation <- apply(dat, 2, mean)
# rowMeans or colMeans to do similar things as apply()

# Subset data example
animal <- c("m", "o", "n", "k", "e", "y")
mon = animal[1:3] # gives us output for 1-3 
key = animal[4:6] # gives us outpout for 4-6 of animal list

# Plotting example
plot(avg_day_inflammation) # visualizes mean inflamm.

# Summary
# Uploading data
# Manipulating data
#   Variables
#   Dataframes (datasets)
#   Numeric vs character
#   Functions and statistics
#   Plotting
```

2. Creating Functions
```{r setup, include=FALSE}
# Defining a Function
fahrenheit_to_kelvin <- function(temp_F) {
  temp_K <- ((temp_F - 32) * (5 / 9)) + 273.15
  return(temp_K)
}

# freezing point of water 32F
fahrenheit_to_kelvin(32) # 273.15 Kelvin

# boiling point of water
fahrenheit_to_kelvin(212) # 373.15

# create other functions such as C <= K, F <= C
kelvin_to_celsius <- function(temp_K) {
  temp_C <- temp_K - 273.15
  return(temp_C)
}

kelvin_to_celsius(fahrenheit_to_kelvin(32.0))
# freezing point of water is 32F which is 0C

dat <- read.csv("data/inflammation-01.csv", FALSE)

dat <- read.csv(header = FALSE, file = "data/inflammation-01.csv")
dat <- read.csv(FALSE, "data/inflammation-01.csv")

center <- function(data, desired = 0) {
  # return a new vector containing the original data centered around the
  # desired value (0 by default).
  # Example: center(c(1, 2, 3), 0) => c(-1, 0, 1)
  new_data <- (data - mean(data)) + desired
  return(new_data)
}

test_data <- c(0, 0, 0, 0)
center(test_data, 3)

names = (Sabah, Katie)

```

3. Analyzing Multiple Data Sets
```{r setup, include=FALSE}
analyze <- function(filename) {
  # Plots the average, min, and max inflammation over time.
  # Input is character string of a csv file.
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  plot(avg_day_inflammation)
  max_day_inflammation <- apply(dat, 2, max)
  plot(max_day_inflammation)
  min_day_inflammation <- apply(dat, 2, min)
  plot(min_day_inflammation)
}

png("test.png")
analyze("data/inflammation-01.csv")
dev.off()

# how to save all three plots at once

# loops
best_practice <- c("Let", "the", "computer", "do", "the", "work") # making a list that spells a sentence

print_words <- function(sentence) {
  print(sentence[1])
  print(sentence[2])
  print(sentence[3])
  print(sentence[4])
  print(sentence[5])
  print(sentence[6])
} # can adjust function to cut out or add words

print_words(best_practice) 

best_practice[-6] # automatically removes 6th word from vector

print_words(best_practice[-6])

# Let's make a looped function (for loop)
print_words <- function(sentence) {
  for (word in sentence) { # variable is word, collection is sentence
    print(word)
  }
}

print_words(best_practice)
print_words(best_practice[-6])

# basic structure of our for loop
# for (variable in collection) {
#   do things with variable
# }

# Read multiple files in one line
list.files(path = "data", pattern = "csv")
list.files(path = "data", pattern = "inflammation")

# plots for all files
filenames <- list.files(path = "data",  
  # Now follows a regular expression that matches:
  pattern = "inflammation-[0-9]{2}.csv",
 #  the standard file extension of comma-separated values
 #  the variable parts (two digits, each between 0 and 9)
#  the static part of the filenames
                        full.names = TRUE)
filenames <- filenames[1:3]
for (f in filenames) {
  print(f)
  analyze(f)
}

# Summary 
# Analyzing of multiple datasets
# Practiced using functions that we created
# For loop we realized we could make function more efficient
# For loop has many uses: make data upload more efficient, and other things
```

4. Making Choices
```{r setup, include=FALSE}

# Re-create analyze function
analyze <- function(filename) {
  # Plots the average, min, and max inflammation over time.
  # Input is character string of a csv file.
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  plot(avg_day_inflammation)
  max_day_inflammation <- apply(dat, 2, max)
  plot(max_day_inflammation)
  min_day_inflammation <- apply(dat, 2, min)
  plot(min_day_inflammation)
}

# Create a nested function of analyze
analyze_all <- function(pattern) {
  # Runs the function analyze for each file in the current working directory
  # that contains the given pattern.
  filenames <- list.files(path = "data", pattern = pattern, full.names = TRUE)
  for (f in filenames) {
    analyze(f)
    }
}

# Save the figures as PDFs
pdf("inflammation-01.pdf") # this saves all of them 
analyze("data/inflammation-01.csv")
dev.off()

png("inflammation-01.png") # this only saves the most recent of 3 figures
analyze("data/inflammation-01.csv")
dev.off() 

# Conditionals and using if else statements
# Create an if else statement
num <- 37
if (num > 100) {
  print("greater")
} else {
  print("not greater")
}

# Setting up an if statement (without else)
num <- 53
if (num > 100) {
  print("num is greater than 100")
}

# If else to return multiple scenarios (more than 2 scenarios)
sign <- function(num) {
  if (num > 0) {
    return(1)
  } else if (num == 0) {
    return(0)
  } else {
    return(-1)
  }
}

sign(-3)
sign(0)
sign(3)

# Plots Based on Data
dat <- read.csv("data/inflammation-01.csv", header = FALSE)
# Write a function
plot_dist <- function(x, threshold) {
  if (length(x) > threshold) {
    boxplot(x)
  } else {
    stripchart(x)
  }
}

plot_dist(dat[, 10], threshold = 10)
plot_dist(dat[1:5, 10], threshold = 10) 


# Automatically create figures within analyze function
analyze <- function(filename, output = NULL) {
  # Plots the average, min, and max inflammation over time.
  # Input:
  #    filename: character string of a csv file
  #    output: character string of pdf file for saving
  if (!is.null(output)) {
    pdf(output)
  }
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  plot(avg_day_inflammation)
  max_day_inflammation <- apply(dat, 2, max)
  plot(max_day_inflammation)
  min_day_inflammation <- apply(dat, 2, min)
  plot(min_day_inflammation)
  if (!is.null(output)) {
    dev.off()
  }
}

output <- NULL
is.null(output)


analyze("data/inflammation-01.csv")
```

5-7. On own time with helpers

8. Making Packages in R
```{r setup, include=FALSE}
install.packages("knitr")
library("knitr")
packageVersion("knitr") # v1.20

# Reasons to not automatically update a script
# Sometimes better because they may have majorly changed script
# You can define the version you want to install 
```

Subset of our data
```{r setup, include=FALSE}
subsetexample = dat[1:5, 10:15] # subset from when we read our data earlier
# read.csv()
write.csv(subsetexample, file = 'subsetexample_cleaned.csv') 
summary(subsetexample) # we can convert to a table output
# If you have a blank cell fill it with NA so that you can execute functions

# Understanding Factors
sex <- factor(c("male", "female", "female", "male"))
levels(sex)
nlevels(sex)

# data.frame (character vs numeric)
# vectors and lists
# elements
# variables (ways of classifying data.frame)
# functions (loops, conditionals)
# factors (ways of classifying data.frame)

# What's the difference between a list and a vector? 
list_example = dat[,10]
str(list) # list you have characters 
# vectors you can only have numbers, must be same data type
# data frame can be mixed

analyze <- function(filename) {
  # Plots the average, min, and max inflammation over time.
  # Input is character string of a csv file.
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  plot(avg_day_inflammation)
  max_day_inflammation <- apply(dat, 2, max)
  plot(max_day_inflammation)
  min_day_inflammation <- apply(dat, 2, min)
  plot(min_day_inflammation)
}

filenames <- list.files(path = "data", pattern = "inflammation-[0-9]{2}.csv", full.names = TRUE)

a <- 1:10
b <- 1:10

res <- numeric(length = length(a))

for 
(i in seq_along(a)) {
  res[i] <- a[i] + b[i]
}

res

sapply(filenames, FUN = analyze)

```



```{r setup, include=FALSE}
library(gglpot2)
library(knitr)
```